\section*{Acknowledgements}

We joke about how thank you speeches are often overlong, but in my turn to write
out acknowledgements I realise the quantity of thank yous is small substitute
for the quality of support. The people below---and others unmentioned---have
been an embarassment of riches. I am fortunate to be able to list their names.

My supervisors
John Baez, for championing my work. For being (thanks Lisa!). John has spent
days of his life answering my emails, let alone skyping me. Throughout my DPhil
I made an annual pilgrimage to see John; . For being patient

Bob Coecke. Bob didn't so much support my application as embody it, showing far
more belief in my abilities and . I would not be

Rob Ghrist, for being ever generous and supportive, and not least for giving me
somewhere to continue this work.

Jamie Vicary supervised my masters' thesis, and didn't really seem to stop. 

My coauthors
John, again
Brandon Coya
Hugo Nava-Kopp
Blake Pollard
Pawe\l\ Soboc\'inski
Paolo Rapisarda

Writing adviser Rashmi Kumar

comments on drafts:
all of the above plus Omar Camarena, Dan Marsden, Bernhard Reinke, Jason Erbele, Stefano
Gogioso,

My fellow mathematicians and friends
Ross Atkins
Fiona Skerman
Kati Cohn-Gordon
Greg Henselman 
Daniel Trewartha

Shiori Shakuto
Bahar Tuncgenc
Mustak Ayub
Adam Berrington


My partner, Stephanie Lin

And my family Mum, Dad, Justin, Calvin, Tania. When I was nine my parents handed
me a three month reading ban. Good sense like that has kept me sane and healthy.

All the organisations: not only the Department of Computer Science and the
University of Oxford, who gave me space to work and, but the Clarendon Fund,
Hertford College, the Queen Elizabeth Scholarships, the Centre for Quantum
Technologies, Singapore, the Department of Mathematics at the University of Pennsylvania


