\begin{thebibliography}{RSW00}

    \bibitem[AC04]{AC04}
    S.\ Abramsky and B.\ Coecke.
    \newblock A categorical semantics of quantum protocols.
    \newblock {\em Proceedings of the 19th IEEE Conference on Logic in
    Computer Science (LiCS'04)}, IEEE Computer Science Press, 2004.
    \newblock \href{http://arxiv.org/abs/quant-ph/0402130}{arXiv:quant-ph/0402130}.

    \bibitem[AHS]{AHS} 
    J.\ Ad\'amek, H.\ Herrlich, G.\ E.\ Strocker. 
    \newblock {\sl Abstract and Concrete Categories: The Joy of Cats}.
    \newblock
    \href{http://katmat.math.uni-bremen.de/acc/acc.pdf}{http://katmat.math.uni-bremen.de/acc/acc.pdf}.


%    \bibitem{ALT} J.\ Alman, C.\ Lian and B.\ Tran, Circular planar electrical
%     networks I: the electrical poset $EP_{n}$. Available as
%      \href{http://arxiv.org/abs/1309.2697}{arXiv:1309.2697}.

    \bibitem[ASW11]{ASW} 
    L.\ Albasini, N.\ Sabadini, R.\ F.\ C.\ Walters. 
    \newblock The compositional construction of Markov processes.
    \newblock \emph{Applied Categorical Structures} {\bf 19} (2011), 425--437. 
    \newblock doi: 10.1007/s10485-010-9233-0.

    \bibitem[Amb91]{Am} 
    S.\ J.\ Ambler. 
    \newblock {\em First order linear logic in symmetric monoidal closed
    categories}. 
    \newblock PhD thesis, University of Edinburgh, 1991.

    \bibitem[AM80]{AM} 
    M.\ A.\ Arbib, E.\ G.\ Manes. 
    \newblock Machines in a category.
    \newblock {\em Journal of Pure and Applied Algebra}, 19:9--20, 1980.


    \bibitem[Ati88]{At} M.\ Atiyah. 
    \newblock Topological quantum field theories.
    \newblock {\em Inst. Hautes \'Etudes Sci. Publ. Math.}, 68:175--186, 1988. 

    \bibitem[AV73]{AV} 
    B.\ D.\ O.\ Anderson and S.\ Vongpanitlerd. 
    \newblock {\sl Network Analysis and Synthesis: a Modern Systems Theory
    Approach}, Prentice-Hall, Englewood Cliffs, 1973.

    \bibitem[Bae06]{Ba1} 
    J.\ C.\ Baez. 
    \newblock Quantum quandaries: a category-theoretic
    perspective. 
    \newblock In {\sl Structural Foundations of Quantum Gravity}, eds.\
    S.\ French, D.\ Rickles and J.\ Saatsi, Oxford U.\ Press, Oxford,
    2006, pp.240--265. 
    \newblock \href{http://arxiv.org/abs/quant-ph/0404040}
    {arXiv:quant-ph/0404040}.

    \bibitem[BE15]{BE} 
    J.\ C.\ Baez and J.\ Erbele.
    \newblock Categories in control. 
    \newblock {\em Theory and Applications of Categories}, {\bf 30}:836--881,
    2015. 
    \newblock Available at
    \href{http://www.tac.mta.ca/tac/volumes/30/24/30-24abs.html}{http://www.tac.mta.ca/tac/volumes/30/24/30-24abs.html}.

    \bibitem[BF]{BF} 
    J.\ C.\ Baez, B.\ Fong. 
    \newblock A compositional framework for passive linear networks.
    \newblock Preprint. 
    \newblock \href{http://arxiv.org/abs/1504.05625}{arXiv:1504.05625}. 

    \bibitem[BFP16]{BFP} 
    J.\ C.\ Baez, B.\ Fong, B.\ Pollard. 
    \newblock A compositional framework for Markov processes.
    \newblock {\sl Journal of Mathematical Physics}, 2016.
    \newblock \href{doi}{doi}

    \bibitem[BL11]{BL} 
    J.\ C.\ Baez and A.\ Lauda.
    \newblock A prehistory of $n$-categorical
    physics. 
    \newblock In {\em Deep Beauty: Mathematical Innovation and
    the Search for an Underlying Intelligibility of the Quantum World}, ed.\
    Hans Halvorson, Cambridge U.\ Press, Cambridge, 2011, pp.13--128.  
    \newblock \href{http://arxiv.org/abs/0908.2469}{arXiv:0908.2469}.

    \bibitem[BS11]{BaezStay} 
    J.\ C.\ Baez and M.\ Stay. 
    \newblock Physics, topology, logic and 
    computation: a Rosetta Stone. 
    \newblock In {\sl New Structures for Physics}, ed.\ B.\ Coecke,  
    Lecture Notes in Physics, vol.\ 813, Springer, Berlin, 2011, pp.\ 173--286.
    \newblock \href{http://arxiv.org/abs/0903.0340}{arXiv:0903.0340}.

    \bibitem[BS82]{BS} 
    P.\ Bamberg and S.\ Sternberg. 
    \newblock {\em A Course of Mathematics for
    Students of Physics}, Vol.\ 2, Chap.\ 12: The theory of electrical circuits,
    Cambridge U.\ Press, Cambridge, 1982. 

    \bibitem[Ben63]{Ben63} 
    J.\ B\'enabou.
    \newblock Cat\'egories avec multiplication. 
    \newblock C. R. Acad. Sci. Paris, 256:1887–1890, 1963.

    \bibitem[Ben67]{Be} 
    J.\ B\'enabou.
    \newblock Introduction to bicategories I.
    \newblock In \textsl{Reports
    of the Midwest Category Seminar}, eds.\ J.\ B\'enabou \textit{et al},
    Springer Lecture Notes in Mathematics {\bf 47}:1--77, New York, 1967.

%	\bibitem{B} S.\ Benenti, {\sl Hamiltonian Structures and Generating
%      Families}, Springer, Berlin, 2011.

    \bibitem[Ber50]{Ber50} 
    L.\ von Bertalanffy. 
    \newblock An Outline of General Systems Theory.
    \newblock {\em British Journal of the Philosophy of Science}, 1:134--165,
    1950.
    
    \bibitem[Ber68]{Ber68} 
    L.\ von Bertalanffy. 
    \newblock {\em General Systems Theory: Foundations, Development,
    Applications},
    George Braziller, New York.
    \newblock Available at
    \href{https://monoskop.org/images/7/77/Von_Bertalanffy_Ludwig_General_System_Theory_1968.pdf}{https://monoskop.org/images/7/77/Von\_Bertalanffy\_Ludwig\_General\_System\_Theory\_1968.pdf}.

    \bibitem[BSZ14]{BSZ} 
    F.\ Bonchi, P.\ Soboci\'nski, F. Zanasi.
    \newblock A categorical semantics of signal flow graphs.
    \newblock In \emph{CONCUR 2014: Concurrency Theory}, eds.\ P.\ Baldan and
    D.\ Gorla, Lecture Notes in Computer Science, 8704:435--450, Springer, 2014.
    \newblock Also available at
    \href{http://users.ecs.soton.ac.uk/ps/papers/sfg.pdf}
    {http://users.ecs.soton.ac.uk/ps/papers/sfg.pdf}.
    
    \bibitem[BSZ15]{BSZ3} 
    F.\ Bonchi, P.\ Soboci\'nski, F.\ Zanasi.
    \newblock Full abstraction for signal flow graphs.
    \newblock In \emph{PoPL '15}, pp 515--526, 2015.

    \bibitem[BSZ16]{BSZ2} 
    F.\ Bonchi, P.\ Soboci\'nski, F.\ Zanasi.
    \newblock Interacting Hopf algebras.
    \newblock {\em J Pure Appl Algebra}, 2016.


   % doi:10.1145/2676726.2676993.
%    \bibitem{Br} J.\ F.\ Broenink, Introduction to physical systems modeling
%     with bond graphs, in {\sl The SiE Whitebook on Simulation Methodologies},
%      1999.

    \bibitem[Bro07]{Brown} 
    F.\ T.\ Brown.
    \newblock \textsl{Engineering System Dynamics: a 
    Unified Graph-Centered Approach}, Taylor and Francis, New York, 2007. 

    \bibitem[Bru31]{Brune} 
    O.\ Brune.
    \newblock Synthesis of a finite two-terminal network whose
    driving-point impedance is a prescribed function of frequency, Doctoral
    thesis, MIT, 1931.
    \newblock Available at
    \href{http://dspace.mit.edu/bitstream/handle/1721.1/10661/36311006.pdf?sequence=1}
    {http://dspace.mit.edu/bitstream/handle/1721.1/10661/36311006.pdf?sequence=1}.

    \bibitem[BG01]{BG} 
    R.\ Bruni, F.\ Gadducci.
    \newblock Some algebraic laws for spans (and their
    connections with multi-relations).
    \newblock In \textsl{RelMiS 2001}, Elsevier, 2001.

    \bibitem[BGM02]{BGM} 
    R.\ Bruni, F.\ Gadducci, U.\ Montanari.
    \newblock Normal forms for algebras of connections.
    \newblock {\em Theoretical Computer Science}, {\bf 286}(2): 247--292, 2002. 
    \newblock \href{http://dx.doi.org/10.1016/S0304-3975(01)00318-8}{doi:10.1016/S0304-3975(01)00318-8}.

    \bibitem[Bud74]{Budak} 
    A.\ Budak.
    \newblock {\sl Passive and Active Network Analysis and 
    Synthesis}, Houghton Mifflin, Boston, 1974.

    \bibitem[Bur93]{Bu} 
    A.\ Burroni.
    \newblock Higher dimensional word problems with
    applications to equational logic.
    \newblock \textsl{Theoretical Computer Science}, {\bf 115}(1):43--62, 1993.
    \newblock \href{http://dx.doi.org/10.1016/0304-3975(93)90054-W}{doi:10.1016/0304-3975(93)90054-W}.


%    \bibitem{C} G.\ E.\ Ching, Topological concepts in networks; an application
%      of homology theory to network analysis, {\sl Proc. 11th. Midwest
%     Conference on Circuit Theory}, University of Notre Dame, 1968, pp.\
%      165--175.

    \bibitem[Car91]{Ca} 
    A.\ Carboni.
    \newblock Matrices, relations and group representations.
    \newblock {\sl J.\ Algebra}, 138:497--529, 1991.

    \bibitem[CW87]{CW}  
    A.\ Carboni, R.\ F.\ C.\ Walters.
    \newblock Cartesian bicategories I.
    \newblock \textsl{J. Pure Appl. Alg.} {\bf 49}:11--32, 1987. 

    \bibitem[CT05]{CT} 
    D.\ Cimasoni, V.\ Turaev.
    \newblock A Lagrangian representation of tangles.
    \newblock {\sl Topology} {\bf 44}:747--767, 2005.
    
    \bibitem[Coe]{Coe} 
    B.\ Coecke.
    \newblock Kindergarten quantum mechanics.
    
    
    \bibitem[CK16]{CK} 
    B.\ Coecke, A.\ Kissinger.
    \newblock \textsl{Picturing Quantum Processes: A First Course in Quantum Theory and Diagrammatic Reasoning}.
    \newblock Cambridge University Press, 2016.

    \bibitem[CP11]{CP} 
    B.\ Coecke, \'E.\ O.\ Paquette.
    \newblock Categories for the practising physicist.
    \newblock In {\sl New Structures for Physics}, ed.\ B.\ Coecke,  
    Lecture Notes in Physics, vol.\ 813, Springer, Berlin, 2011, pp.\ 173--286.
    \newblock \href{http://arxiv.org/abs/0905.3010}{arXiv:0905.3010}.

    \bibitem[CPP09]{CPP} 
    B.\ Coecke, E.\ Paquette, D.\ Pavlovic.
    \newblock Quantum and classical structuralism.
    \newblock In \textsl{Semantic Techniques for Quantum
    Computation}, eds. I.\ Mackie and S.\ Gay, Cambridge University Press, 2009,
    pp.29--69.
    \newblock \href{http://arxiv.org/abs/0904.1997}{arXiv:0904.1997}.

    \bibitem[Cou16]{Cou16} 
    K.\ Courser.
    \newblock A bicategory of decorated cospans.
    \newblock Preprint. 
    \newblock \href{http://arxiv.org/abs/1606.}{arXiv:1606.}.

    \bibitem[CF16]{CF}
    B.\ Coya, B.\ Fong.
    \newblock Corelations are the prop for extraspecial commutative Frobenius
    monoids.
    \newblock Preprint.
    \newblock \href{http://arxiv.org/abs/1601.02307}{arXiv:1601.02307}.

    \bibitem[Dav80]{D} 
    B.\ Davies.
    \newblock A web of naked fancies?.
    \newblock {\sl Phys. Educ.} {\bf 15}:57--61, 1980.

    \bibitem[Deu0]{Deu}
    D.\ Deutch
    \newblock Constructor theory
    
    \bibitem[Dij]{Dij}
    E.\ W.\ Dijkstra
    \newblock On the cruelty of really teaching comp
    \newblock here.


    \bibitem[DS84]{DS} 
    P.\ G.\ Doyle, J.\ L.\ Snell.
    \newblock {\sl Random Walks and Electrical Circuits}, Mathematical Association of America, 1984.
    \newblock Also available at \href{http://www.math.dartmouth.edu/~doyle/}
    {http://www.math.dartmouth.edu/\~{}doyle/}.

    \bibitem[DP03]{DP1} 
    K.\ Do\v{s}en, Z.\ Petri\'c.
    \newblock Generality of proofs and its Brauerian representation.
    \newblock \textsl{J. Symbolic Logic} {\bf 68}(3):740--750, 2003.
    \newblock
    \href{http://doi.org/10.2178/jsl/1058448435}{doi:10.2178/jsl/1058448435}.

%  \bibitem[DP13]{DP2} K.\ Do\v{s}en and Z.\ Petri\'c, Syntax for split preorders,
%    \textsl{Annals of Pure and Applied Logic} {\bf 164}(4):443--481, 2013.
%    \href{http://doi.org/10.1016/j.apal.2012.10.008}{doi:10.1016/j.apal.2012.10.008}. 
  \bibitem[Ell14]{El} 
  D.\ Ellerman.
  \newblock An introduction to partition logic.
  \newblock \textsl{Logic Journal of the Interest Group in Pure and Applied
  Logic}, {\bf 22}(1):94--125, 2014.
  \newblock
  \href{http://doi.org/10.1093/jigpal/jzt036}{doi:10.1093/jigpal/jzt036}.
    
  \bibitem[Erb16]{Erb16}
    J.\ Erbele.
    \newblock Categories in Control: PROPs .
    \newblock PhD thesis, University of California, Riverside, 2016. 

    \bibitem[Fon15]{Fon15}
    B.\ Fong.
    \newblock Decorated Cospans.
    \newblock {\em Theory and Applications of Categories}, 30(33):25--50, 2015.
    \newblock \href{http://arxiv.org/abs/1502.00872}{arXiv: 1502.00872}.

    \bibitem[FRS16]{FRS16}
    B.\ Fong, P.\ Rapisarda, P.\ Sobocinski.
    \newblock A categorical approach to open and interconnected dynamical
    systems.
    \newblock To appear in {\sl Proceedings of Logic in Computer Science 2016,
    LICS16}.
    \newblock \href{http://arxiv.org/abs/1510.05076}{arXiv:1510.05076}.

    \bibitem[For61]{Fo} 
    J.\ W.\ Forrester.
    \newblock {\sl Industrial Dynamics}, Pegasus Communications, 1961. 

%    \bibitem{Fr} B.\ Friedland, {\sl Control System Design: An Introduction to
%      State-Space Methods}, McGraw-Hill, New York, 1986.

    \bibitem[FH15]{FH}
    P.A.~Fuhrmann and U. Helmke.
    \newblock \emph{The Mathematics of Networks of Linear Systems},
    Springer-Verlag, 2015. 

    \bibitem[Fuk80]{Fukushima}
    M.\ Fukushima.
    \newblock {\sl Dirichlet Forms and Markov Processes},
    North-Holland, Amsterdam, 1980.

    \bibitem[Gog75]{Go}
    J.\ Goguen.
    \newblock Discrete-time machines in closed monoidal categories I.
    \newblock \emph{J Comput Systems Sci}, 10(1):1--43, 1975.
    \newblock
    \href{http://doi.org/10.1016/S0022-0000(75)80012-2}{doi:10.1016/S0022-0000(75)80012-2}.

    \bibitem[Gra01]{Gr} 
    M.\ Grandis.
    \newblock Finite sets and symmetric simplicial sets.
    \newblock \textsl{Theory Appl. Categ.} {\bf 8}:244--252, 2001.
    \newblock Available at
    \href{http://www.tac.mta.ca/tac/volumes/8/n8/8-08abs.html}
    {http://www.tac.mta.ca/tac/volumes/8/n8/8-08abs.html}.

%    \bibitem{GK} P.\ W.\ Gross and P.\ R.\ Kotiuga, {\sl Electromagnetic Theory
%      and Computation: A Topological Approach}, Cambridge U.\ Press, Cambridge,
%      2004. 

    \bibitem[Har23]{H}
    I.\ B.\ Hart.
    \newblock {\sl Makers of Science}, Oxford U.\ Press,
    Oxford, 1923, p.\ 243.

%    \bibitem{Hu} J.\ H.\ Hubbard, The Bott--Duffin synthesis of electrical
%      circuits, in {\sl A Celebration of the Mathematical Legacy of Raoul Bott},
%     ed.\ P.\ R.\ Kotiuga, CRM Proceedings and Lecture Notes, American
%      Mathematical Society, 2010.
    
    \bibitem[Har15]{Har15}
    P.\ Harron.
    \newblock .
    \newblock PhD thesis, Princeton University, 2015. 
    
    \bibitem[Jan86]{Jan86}
    T.\ Janssen
    \newblock {\em Foundations and applications of Montague grammar. Part 1:
    Philosohpy, framework, computer science}, CWI tract 19, Centrum voor
    Wiskunde en Informatica, 1986.
    \newblock
    \href{http://www.illc.uva.nl/Research/Publications/Dissertations/HDS-10-Theo-Janssen.text.pdf}{http://www.illc.uva.nl/Research/Publications/Dissertations/HDS-10-Theo-Janssen.text.pdf}.
    
    \bibitem[JW00]{JW}
    R.\ Jayawardene and O.\ Wyler.
    \newblock Categories of relations and functional relations.
    \newblock \emph{Appl Categorical Str} 8(1):279--305, 2000.
    
    \bibitem[Jek]{Jek}
    D.\ Jekel.
    \newblock Layering $\partial$-graphs and networks. 
    \newblock Preprint, 2016.
    \newblock \href{http://arxiv.org/abs/1601.00247}{arXiv:1601.00247}

    \bibitem[JS91]{JS91} 
    A.\ Joyal, R.\ Street.
    \newblock The geometry of tensor calculus I.
    \newblock \emph{Advances in Mathematics} {\bf 88}(1):55--112, 1991.
    \newblock \href{http://doi.org/10.1016/0001-8708(91)90003-P}{doi:10.1016/0001-8708(91)90003-P}.

    \bibitem[JS93]{JS93} 
    A.\ Joyal, R.\ Street.
    \newblock Braided tensor categories.
    \newblock {\sl Adv.\ Math.}, {\bf 102}:20--78, 1993.

    \bibitem[Kai80]{Ka} 
    T.\ Kailath.
    \newblock \emph{Linear Systems}, Prentice-Hall, New Jersey, 1980.

    \bibitem[KSW97]{KSW}
    P.\ Katis, N.\ Sabadini, R.\ F.\ C.\ Walters.
    \newblock Span(Graph): A Categorial Algebra of Transition Systems.
    \newblock {\sl AMAST '97 Proceedings of the 6th International Conference on
    Algebraic Methodology and Software Technology}, 307--321, 1997. 

    \bibitem[KSW00]{KSW2} 
    P.\ Katis, N.\ Sabadini, R.\ F.\ C.\ Walters.
    \newblock On the algebra of systems with feedback \& boundary.
    \newblock {\sl Rendiconti del Circolo Matematico di Palermo Serie II},
    Suppl. {\bf 63}:123--156, 2000.

%    \bibitem{KT} H.\ W.\ Knobloch and M.\ Thoma, Aspects of modern control
%      theory, in {\sl Perspectives in Mathematics: Anniversary of Oberwolfach
%      1984}, Birkh\"auser, Basel, 1984, pp.\ 337--376.

%    \bibitem{K} J.\ Kigami, {\sl Analysis on Fractals}, Cambridge U. Press, Cambridge, 
%	2008.
%     First 60 pages available at
%      \href{http://www-an.acs.i.kyoto-u.ac.jp/~kigami/AOF.pdf}{http://www-an.acs.i.kyoto-u.ac.jp/\~{}kigami/AOF.pdf}.
%   I think some nice results on Dirichlet forms are proved here.

    \bibitem[KMR90]{KRM} 
    D.\ C.\ Karnopp, D.\ L.\ Margolis, R.\ C.\ Rosenberg.
    \newblock \textsl{System Dynamics: a Unified Approach}, Wiley, New York, 1990.

    \bibitem[Kis14]{Ki}
    A.\ Kissinger.
    \newblock Finite matrices are complete for (dagger-)hypergraph categories.
    \newblock preprint.
    \newblock \href{http://arxiv.org/abs/1406.5942}{arXiv:1406.5942}.

    \bibitem[Lac04]{La} 
    S.\ Lack.
    \newblock Composing PROPs.
    \newblock {\em Theory and Applications of Categories}, {\bf 13}(9):147--163,
    2004.
%    \newblock Available at
%    \href{http://www.tac.mta.ca/tac/volumes/13/9/13-09abs.html}
%    {http://www.tac.mta.ca/tac/volumes/13/9/13-09abs.html}.

    \bibitem[Law63]{Law} 
    F.\ W.\ Lawvere.
    \newblock Functorial semantics of algebraic theories and
    some algebraic problems in the context of functorial semantics of algebraic
    theories.
    \newblock \textsl{Reprints in Theory Appl.\ Categ.} {\bf 5}:1--121.
    \newblock 
    Available at
    \href{http://www.tac.mta.ca/tac/reprints/articles/5/tr5abs.html}
    {http://www.tac.mta.ca/tac/reprints/articles/5/tr5abs.html}.

    \bibitem[LR03]{LR} 
    F.\ W.\ Lawvere, R.\ Rosebrugh.
    \newblock \emph{Sets for Mathematics},
    Cambridge University Press, 2003.

    \bibitem[LO73]{vLO} 
    M.\ van Lier, R.\ Otten.
    \newblock Planarization by transformation.
    \newblock {\sl IEEE Transactions on Circuit Theory} {\bf 20}:169--171, 1973.

    \bibitem[MR91]{MR}
    Z.-M.\ Ma and M.\ R\"ockner.
    \newblock {\sl Introduction to the Theory of
    (Non-Symmetric) Dirichlet Forms}, Springer, Berlin, 1991.

    \bibitem[Mac63]{Mac63} 
    S.\ Mac Lane.
    \newblock Natural associativity and commutativity.
    \newblock {\sl Rice Univ. Studies} {\bf 49}:28--46, 1963. 

    \bibitem[Mac65]{Mac65}
    S.\ Mac\ Lane.
    \newblock Categorical algebra.
    \newblock \emph{Bull. Amer. Math. Soc.} {\bf 71}(1):40--106, 1965.
    \newblock 
    \href{http://doi.org/10.1090/S0002-9904-1965-11234-4}{doi:10.1090/S0002-9904-1965-11234-4}.

    \bibitem[Mac98]{Mac98}
    S.\ Mac Lane.
    \newblock {\sl Categories for the Working Mathematician},
    Springer, 1998.
    
    \bibitem[Mil00]{Mi}
    S.\ Milius.
    \newblock Relations in categories, MA thesis, York University,
    Ontario, 2000.

    \bibitem[Mon70]{Mon70}
    R.\ Montague.
    \newblock Universal Grammar.
    \newblock Reprinted in R. Thomason (ed.), {\em Formal Philosophy}, Yale
    University Press, 222--246, 1974. 


    \bibitem[Mor]{Mo} J.\ Morton, 
    
    \bibitem[Nob06]{Nob06}
    D.\ Noble.
    \newblock {\em The Music of Life: Biology Beyond the Genome}, Oxford, Oxford
    University Press, 2006.

    \bibitem[Obe90]{Ob}
    U.\ Oberst.
    \newblock Multidimensional constant linear systems.
    \newblock \emph{Acta Applicandae Mathematicae}, vol. 20, pp 1--175, 1990.

    \bibitem[Odu84]{Od} 
    H.\ T.\ Odum.
    \newblock {\sl Ecological and General Systems: An
    Introduction to Systems Ecology}, Wiley, New York, 1984.

    \bibitem[Ohm27]{O} 
    G.\ Ohm.
    \newblock {\sl Die Galvanische Kette, Mathematisch Bearbeitet},
    T. H. Riemann, Berlin, 1827.
    \newblock Also available at
    \href{http://www.ohm-hochschule.de/bib/textarchiv/Ohm.Die_galvanische_Kette.pdf}{www.ohm-hochschule.de/bib/textarchiv/Ohm.Die\_galvanische\_Kette.pdf}.

    \bibitem[OFM07]{OFM} 
    R.\ Olfati-Saber, J.A.\ Fax, R.M.\ Murray.
    \newblock Consensus and cooperation in networked multi-agent systems.
    \newblock \emph{Proc.\ IEEE}, 96(1):215--233, 2007.

    \bibitem[Ols43]{Ol} 
    H.\ F.\ Olson.
    \newblock {\sl Dynamical Analogies}, Van
    Nostrand, New York, 1943.
    \newblock Also available at 
    \href{https://archive.org/details/DynamicalAnalogies}
    {https://archive.org/details/DynamicalAnalogies}.

%    \bibitem{PW} J.\ W.\ Polderman, J.\ Willems, {\sl Introduction to the
%      Mathematical Theory of Systems and Control}, Springer, Berlin, 1998. 

    \bibitem[Pay61]{Pa}
    H.\ M.\ Paynter.
    \newblock {\sl Analysis and Design of Engineering
    Systems}, MIT Press, Cambridge, Massachusetts, 1961.

%    \bibitem{Pe} R. Penrose. Applications of negative dimensional tensors, in
%      {\sl Combinatorial Mathematics and its Applications}, ed.\ D.\ J.\ A.\
%      Welsh, Academic Press, New York, 1971, 221--244. 

    \bibitem[PT08]{PT}
    P.\ Piccione and D.\ V.\ Tausk.
    \newblock A student's guide to symplectic
    spaces, Grassmannians and Maslov index, Publicac c\~oes Matem\'aticas do
    IMPA, Rio de Janeiro, 2008.
    \newblock Also available at
    \href{http://www.ime.usp.br/~piccione/Downloads/MaslovBook.pdf}
    {http://www.ime.usp.br/$\sim$piccione/Downloads/MaslovBook.pdf}.

    \bibitem[Pir02]{Pi}
    T.\ Pirashvili.
    \newblock On the PROP corresponding to bialgebras.
    \newblock 
    \emph{Cah. Top. G\'{e}om. Diff. Cat}, {\bf 43}(3):221--239, 2002.
    \newblock 
    \href{http://arxiv.org/abs/math/0110014}{arXiv:math/0110014}.

    \bibitem[Pol16]{Po}
    B.\ Pollard.
    \newblock Open Markov processes: a compositional
    perspective on non-equilibrium steady states in biology.
    \newblock {\sl Entropy}.

    \bibitem[RSW05]{RSW05} 
    R.\ Rosebrugh, N.\ Sabadini, R.\ F.\ C.\ Walters.
    \newblock Generic
    commutative separable algebras and cospans of graphs.
    \newblock \textsl{Th.\ App.\
    Cat.\ }\textbf{15} (2005), 264--277.
    \newblock Available at
    \href{http://www.tac.mta.ca/tac/volumes/15/6/15-06abs.html}{http://www.tac.mta.ca/tac/volumes/15/6/15-06abs.html}.

    \bibitem[RSW08]{RSW08} 
    R.\ Rosebrugh, N.\ Sabadini, R.\ F.\ C. Walters.
    \newblock Calculating
    colimits compositionally.
    \newblock In \emph{Montanari Festschrift, LNCS} {\bf 5065}:581--592, 2008.

    \bibitem[Ros70]{Ro} 
    H.\ H.\ Rosenbrock.
    \newblock \emph{State Space and Multivariable Theory},
    Wiley, New York, 1970. 

%    \bibitem{R} J.\ P.\ Roth, Existence and uniqueness of solutions to
%     electrical network problems via homology sequences, {\sl Mathematical
%     Aspects of Electrical Network Analysis}, eds.\ H.\ S.\ Wilf and F.\ Harary, 
%     AMS, Providence, Rhode Island, 1971, pp.\ 113--118. 

    \bibitem[Sab97]{Sabot1997}
    C.\ Sabot.
    \newblock Existence and uniqueness of diffusions on
    finitely ramified self-similar fractals, Section 1: Dirichlet forms on
    finite sets and electrical networks.
    \newblock {\sl Annales Scientifiques de l'\'Ecole
    Normale Sup\'erieure}, S\'er. 4({\bf 30}):605--673, 1997.
    \newblock Also available at
    \href{http://www.numdam.org/numdam-bin/item?id=ASENS_1997_4_30_5_605_0}{http://www.numdam.org/numdam-bin/item?id=ASENS\_1997\_4\_30\_5\_605\_0}.

    \bibitem[Sab04]{Sabot2004} 
    C.\ Sabot.
    \newblock Electrical networks, symplectic reductions, and
    application to the renormalization map of self-similar lattices.
    \newblock {\sl
    Proc.\ Symp.\ Pure Math.} {\bf 72}:155--205, 2004.
    \newblock Also available as
    \href{http://arxiv.org/abs/math-ph/0304015}{arXiv:math-ph/0304015}.

    \bibitem[Sch88]{Sc} 
    J.\ M.\ Schumacher.
    \newblock Transformations of linear systems under
    external equivalence.
    \newblock \emph{Linear Algebra and its Applications} 102,
    pp 1--33, 1988.

    \bibitem[SS71]{SS71} 
    D.\ Scott, C.\ Strachey.
    \newblock {\em Toward a mathematical semantics for computer languages},
    Oxford Programming Research Group Technical Monograph PRG-6, 1971.


    \bibitem[Sel07]{Sel07} 
    P.\ Selinger.
    \newblock Dagger compact closed categories and completely
    positive maps.
    \newblock In {\sl Proceedings of the 3rd International Workshop on
    Quantum Programming Languages} (QPL 2005), {\sl ENTCS} {\bf 170}:
    139--163, 2007.
    \newblock Also available at
    \href{http://www.mscs.dal.ca/~selinger/papers/dagger.pdf}{http://www.mscs.dal.ca/\~{}selinger/papers/dagger.pdf}.

    \bibitem[Sel11]{Sel11} 
    P.\ Selinger.
    \newblock A survey of graphical languages
    for monoidal categories.
    \newblock In B.\ Coecke, ed., \emph{New Structures for
    Physics}, Lecture Notes in Physics 813:289--355, Springer, 2011.
    \newblock Available
    as \href{http://arxiv.org/abs/0908.3347}{arXiv:0908.3347}.

    \bibitem[Sha42]{Sh}
    C.\ E.\ Shannon.
    \newblock The theory and design of linear differential equation machines, National Defence Research Council, 1942.


    \bibitem[Sle68]{Slepian}
    P.\ Slepian.
    \newblock {\sl Mathematical Foundations of Network
    Analysis}, Springer, Berlin, 1968. 

    \bibitem[Sma72]{Smale}
    S.\ Smale.
    \newblock On the mathematical foundations of electrical
    network theory.
    \newblock {\sl J.\ Diff.\ Geom.} {\bf 7} (1972), 193--210. 

%    \bibitem{So} E.\ D.\ Sontag, {\sl Mathematical Control Theory: Deterministic
%      Finite Dimensional Systems}, Springer, Berlin, 1998. Also available at \\
%     \href{http://www.math.rutgers.edu/~sontag/FTP_DIR/sontag_mathematical_control_theory_springer98.pdf}
%{http://www.math.rutgers.edu/$\sim$sontag/FTP\_DIR/sontag\_mathematical\_control\_theory\\\_springer98.pdf}.

%    \bibitem{So2} E.\ D.\ Sontag, Some new directions in control theory inspired
%      by systems biology, {\sl Syst. Biol.} {\bf 1}(1) (2004), 9-18. Available
%     at
%      \href{http://popmodel.narod.ru/sb/9.pdf}{http://popmodel.narod.ru/sb/9.pdf}.


    \bibitem[Spi13]{Sp} 
    D.\ I.\  Spivak.
    \newblock The operad of wiring diagrams:
    formalizing a graphical language for databases, recursion, and plug-and-play
    circuits.
    \newblock preprint.
    \newblock Available as
    \href{http://arxiv.org/abs/1305.0297}{arXiv:1305.0297}.

    \bibitem[SSR15]{SSR} 
    D.\ I.\ Spivak, P.\ Schultz, D.\ Rupel.
    \newblock String diagrams for traced and compact categories are oriented 1-cobordisms.

    \bibitem[Sta16]{Sta16} 
    M.\ Stay.
    \newblock Compact closed bicategories.
    \newblock {\em Theory and Applications of Categories}, 2013. 
    \newblock \href{http://arxiv.org/abs/1301.1053}{arXiv:1301.1053}.

    \bibitem[Sza13]{Sza13}
    Szab\'o, Zolt\'an Gendler.
    \newblock "Compositionality", The Stanford Encyclopedia of Philosophy
    (Fall 2013 Edition), Edward N. Zalta (ed.).
    \newblock
    http://plato.stanford.edu/archives/fall2013/entries/compositionality/.

%    \bibitem{Th} J.\ U.\ Thoma, {\sl Introduction to Bond Graphs and Their
%      Applications}, Pergamon Press, Oxford, 1975.

    \bibitem[VSL15]{VSL}
    D.\ Vagner, D.\ I.\ Spivak, E.\ Lerman.
    \newblock Algebras of open
    dynamical systems on the operad of wiring diagrams.
    \newblock \emph{Theory and
    Applications of Categories}, vol. 30, no. 51, pp 1793--1822, 2015. 


%   \bibitem{V1} Y.\ C.\ de Verdiere, Th\'eor\`eme de Kirchhoff et th\'eorie de
%      Hodge, {\sl S\'eminaire de Th\'eorie spectrale et g\'eom\'etrie} {\bf 9}
%      (1990-1991), 89--94.

%    \bibitem{V} Y.\ C.\ de Verdiere, Reseaux electriques planaires I, {\sl
%      Comment.\ Math.\ Helv.} {\bf 69} (1994), 351-374.  Also available at
%    \href{http://www-fourier.ujf-grenoble.fr/~ycolver/All-Articles/94a.pdf}
%	{http://www-fourier.ujf-grenoble.fr/\~{}ycolver/All-Articles/94a.pdf}.

%   \bibitem{dVGV} Y.\ C.\ de Verdiere, I.\ Gitler, and D.\ Vertigan, R\'eseaux
%      \'electriques planaire II, {\sl Comment.\ Math.\ Helv.} {\bf 71} (1996),
%      144--167.

    \bibitem[WW15]{WW} S.\ J.\ Wadsley, N.\ Woods. 
    \newblock PROPs for linear systems.
    \newblock preprint. 
    \newblock \href{http://arxiv.org/abs/1505.00048}{arXiv:1505.00048}.

    \bibitem[Wei81]{Weinstein} 
    A.\ Weinstein.
    \newblock Symplectic geometry.
    \newblock {\sl Bull.\ Amer.\ Math.\ 
    Soc.} {\bf 5} (1981), 1--13.
    \newblock Also available at 	
    \href{https://projecteuclid.org/euclid.bams/1183548217}
    {https://projecteuclid.org/euclid.bams/1183548217}.

    \bibitem[Wey23]{Weyl} 
    H.\ Weyl.
    \newblock Repartici\'on de corriente en una red conductora.
    \newblock {\sl Rev. Mat. Hisp. Amer.} {\bf 5}:153--164, 1923.
    \newblock \href{http://math.ucr.edu/home/baez/weyl1923.pdf}{http://math.ucr.edu/home/baez/weyl1923.pdf}


    \bibitem[Wig60]{Wig60}
    E.\ P.\ Wigner.
    \newblock The unreasonable effectiveness of mathematics in the natural
    sciences.
    \newblock {\em Communications on Pure and Applied Mathematics}, 13: 1–14,
    1960.

    \bibitem[Wil86]{Wi3} 
    J.\ C.\ Willems.
    \newblock From time series to linear system---Part I.
    Finite dimensional linear time invariant systems.
    \newblock \emph{Automatica}, vol.  22, no. 5, pp 561--580, 1986. 

    \bibitem[Wil89]{Wi2} 
    J.\ C.\ Willems.
    \newblock Models for dynamics.
    \newblock \emph{Dynamics Reported,
    Volume 2}, eds. U.\ Kirchgraber and H.\ O.\ Walther, John Wiley \& Sons
    Ltd and B.\ G.\ Teubner, 1989.

    \bibitem[Wil07]{Wi} 
    J.\ Willems.
    \newblock The behavioral approach to open and interconnected
    systems.
    \newblock {\sl IEEE Control Systems Magazine} {\bf 27} (2007), 46--99.
    \newblock Also available at
    \href{http://homes.esat.kuleuven.be/~jwillems/}  
    {http://homes.esat.kuleuven.be/\~{}jwillems/}.

    \bibitem[Wil07b]{Wi4} 
    J.\ C.\ Willems.
    \newblock In control, almost from the beginning until the day
    after tomorrow.
    \newblock {\sl European Journal of Control} {\bf 13}:71--81, 2007. 

    \bibitem[Zan15]{Za} F.\ Zanasi. 
    \newblock \emph{Interacting Hopf algebras---the theory of linear
    systems}. 
    \newblock PhD thesis, \'Ecole Normale Sup\'eriere de Lyon, 2015. 
    \newblock Available
    at \href{http://zanasi.com/fabio/IHthesis_FZ.pdf}
    {http://zanasi.com/fabio/IHthesis\_FZ.pdf}.

\end{thebibliography}


%\bibitem[]{}
%.
%\newblock .
%\newblock {\em }, ():, .
%\newblock \href{}{}.
